---
title: "Lecturer Questions: Data summary"
author: "George Kinnear"
date: "17/11/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
theme_set(theme_minimal())
```

```{r read-data, message=FALSE, warning=FALSE}
all_qs_all_lecs <- read_csv("data/ANON_LQ_Maths_All.csv")
anon_courses <- read_csv("data/ANON_courses.csv")

codes <- c("A", "C", "E", "F", "N", "O", "P", "U", "W", "X")
# Collapsing C, P, X, O into Other
collapsed_codes <- c("A", "O", "E", "F", "N", "O", "O", "U", "W", "O")

code_names <- c("Admin", "Convention", "Evaluation", "Fact", "Next step", "Other",
                "Proof framework", "Understanding check", "Warrant", "Example generation")

code_details <- tibble(code = codes, code_name = code_names) %>% 
  mutate(
    collapsed_code = case_when(
      code %in% c("C", "P", "X", "O") ~ "Other",
      TRUE ~ code_name
    )
  ) %>% 
  add_row(
    code = "Total",
    code_name = "Total",
    collapsed_code = "Total"
  )

```
```{r}
# Making list of all lectures and category code combos
lec_cat <- expand_grid(
  all_qs_all_lecs %>% select(course, session) %>% distinct(),
  category = codes
)
```

# Summary of the data

Codes used to categorise questions:

```{r}
code_details %>% 
  basic_kable()
```

## Summary of courses

```{r}
course_demographics <- anon_courses %>% 
  left_join(all_qs_all_lecs %>% select(course, session) %>% distinct(), by = c("anon_code" = "course")) %>% 
  group_by(anon_course) %>% 
  summarise(
    course_level = first(course_level),
    course_topic = first(course_topic),
    room_capacity = max(room_cap),
    num_lecturers = n_distinct(anon_lec),
    num_lectures = n(),
    .groups = "drop"
  ) %>% 
  separate(course_level, into = c("M_or_E", "year"), sep = 1)  %>% 
  select(
    year,
    course_code = anon_course,
    course_topic,
    room_capacity,
    num_lecturers,
    num_lectures
  ) %>% 
  mutate(
    room_capacity_approx = round(room_capacity/25, digits = 0)*25
  ) %>% 
  relocate(room_capacity_approx, .after = room_capacity) %>% 
  arrange(year, -room_capacity, -num_lectures, -num_lecturers, course_topic)

course_demographics %>% 
  select(-room_capacity) %>% 
  rename(
    "Year" = year,
    "Course code" = course_code,
    "Topic" = course_topic,
    "Room capacity (approx.)" = room_capacity_approx,
    "Number of lecturers" = num_lecturers,
    "Number of lectures" = num_lectures
  ) %>% 
  adorn_totals(c("row")) %>% 
  kable(booktabs = T) %>% 
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "top")
  
```

## Courses, lecturers, and question counts
```{r}
all_qs_all_lecs %>% 
  group_by(course) %>% 
  summarise(
    #num_qs = n(),
    num_qs = sum(!is.na(category)), # to count only rows with a question identified
    num_lectures = n_distinct(session),
    .groups = "drop"
  ) %>% 
  separate(course, c("course", "lecturer")) %>% 
  adorn_totals() %>% 
  basic_kable(booktabs = T)
```


## Questions by category
```{r}
all_qs_all_lecs %>% 
  tabyl(course, category) %>% 
  select(-"NA_") %>% 
  adorn_totals(c("row", "col")) %>% 
  basic_kable(booktabs = T)
```


Number of questions that are mathematical, i.e. not Admin or Understanding checks:
```{r}
## Total number of questions
all_qs_all_lecs %>%
  #filter(category != "A" & category != "U") %>% 
  tabyl(course, category) %>% 
  select(-"NA_", -A, -U) %>% 
  adorn_totals(c("row", "col")) %>% 
  basic_kable(booktabs = T)
```


# Questions by category and lecture

```{r}

count_qs_by_cat = all_qs_all_lecs %>% 
  mutate(
    lecture = paste(course, session)
  ) %>% 
  tabyl(lecture, category) %>% 
  select(-"NA_") %>% 
  separate(
    col = lecture,
    into = c("course", "session"),
    sep = " "
  ) %>% 
  adorn_totals(c("col"))

count_qs_by_cat %>% 
  arrange(course, session)
```

# Opportunities to participate

We create a table that counts all:

* questions,
* questions with student responses, and
* participation opportunities

in each of the lectures in the sample.

```{r}
counts_by_lecture <- all_qs_all_lecs %>%
  filter(!is.na(category)) %>% 
  group_by(course, session, category) %>%
  summarise(
    num_q = sum(!is.na(category)),
    num_sr = sum(student_response == "Y"),
    num_po = sum(student_response == "Y" | wait_time >= 3),
    .groups = "drop"
  ) %>%
  # make sure lectures with no questions are included, with counts of 0
  full_join(lec_cat, by = c("course", "session", "category")) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  # tidy up the columns, including replacing category codes with names
  arrange(course, session, category) %>% 
  left_join(code_details, by = c("category" = "code")) %>% 
  select(code_name, everything(), -collapsed_code, -category)
```

The following sections present summaries of this data, with different choices about which categories to include.

Note that the histograms that are plotted show **only the nonzero counts** (to avoid having a large spike at 0 in many cases).

## Summary - all questions

```{r}
# Add totals for each course/session combo
# NB - this was a possible alternative approach
# https://community.rstudio.com/t/getting-summary-by-group-and-overall-using-tidyverse/54656/6

add_group_totals <- function(df) {
  df %>% 
    # Add rows containing the totals
    bind_rows(
      df %>%
        group_by(course, session) %>%
        summarise(across(contains("num_"), sum), .groups = "drop_last") %>% 
        mutate(code_name = "Total")
    ) %>% 
    # Put the totals at the end of each group (just for neatness)
    mutate(is_total = code_name=="Total") %>% 
    arrange(course, session, is_total, code_name) %>% 
    select(-is_total)
}

counts_by_lecture_inc_totals <- add_group_totals(counts_by_lecture)

counts_by_lecture_inc_totals %>% 
  write_csv("data/output-01-counts_by_lecture_inc_totals.csv")
```

```{r}
produce_summary_table <- function(df) {
  df %>%
    group_by(code_name) %>%
    summarise(across(contains("num_"), list(mean = mean, sd = sd)), .groups = "drop_last")%>% 
    # Order the rows by frequency, with total at the end
    mutate(is_total = code_name=="Total") %>% 
    arrange(is_total, -num_q_mean) %>% 
    select(-is_total) %>%
    # Add the 'percentage of group' columns
    mutate(
      "num_sr_mean_pc" = num_sr_mean / num_q_mean * 100,
      "num_po_mean_pc" = num_po_mean / num_q_mean * 100,
      across(contains("pc"), round, digits = 0),
      across(where(is.numeric), round, digits = 2)
    )
}

counts_by_lecture_summary_table <- produce_summary_table(counts_by_lecture_inc_totals)

show_summary_table <- function(counts_by_lecture_summary_table) {
  # This is the unformatted table
  counts_by_lecture_summary_table %>% 
    kable(booktabs = T) %>% 
    kable_styling(full_width = T) %>%
    row_spec(0, angle = 90)
  
  counts_by_lecture_summary_table %>% 
    transmute(
      "Question category" = code_name,
      "Mean number of questions per lecture (s.d.)" = paste0(num_q_mean, " (", num_q_sd, ")"),
      "Mean number of student responses per lecture (as % of category)" = paste0(num_sr_mean, " (", num_sr_mean_pc,"%)"),
      "Mean number of participation opportunities per lecture (as % of category)" = paste0(num_po_mean, " (", num_po_mean_pc,"%)")
    ) %>% 
    kable(booktabs = T) %>% 
    kable_styling(full_width = T)
}

show_summary_table(counts_by_lecture_summary_table)

plot_summary <- function(counts_by_lecture_inc_totals) {

counts_by_lecture_inc_totals %>% 
  mutate(
    code_name = fct_reorder(code_name, num_q, .fun = sum)
  ) %>% 
  pivot_longer(
    cols = contains("num_"),
    names_to = "type",
    values_to = "number"
  ) %>% 
  mutate(
    type = case_when(
      type == "num_q" ~ "questions",
      type == "num_sr" ~ "questions with student response",
      type == "num_po" ~ "participation opportunities"
    )
  ) %>% 
  # This is used to omit cases where the count is 0, to avoid overly spiky histograms
  filter(number > 0) %>% 
  ggplot(aes(x = number)) +
  facet_grid(type ~ code_name) +
  geom_histogram(binwidth = 5) +
  theme_minimal()

}

counts_by_lecture_inc_totals %>% 
  #filter(code_name %in% c("Proof framework")) %>% 
  plot_summary()
```

## Summary - mathematical questions only

Here we remove the `Admin` and `Understanding check` categories.

```{r}

counts_by_lecture_inc_totals_noAU <- counts_by_lecture %>% 
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  add_group_totals()

counts_by_lecture_summary_table_noAU <- produce_summary_table(counts_by_lecture_inc_totals_noAU)

counts_by_lecture_summary_table_noAU %>% 
  show_summary_table()
```

Version of this table with grouped columns, and corresponding LaTeX code for the manuscript:

```{r}
counts_by_lecture_summary_table_noAU %>% 
    select(
      code_name,
      num_q_mean,
      num_q_sd,
      num_sr_mean,
      num_sr_mean_pc,
      num_po_mean,
      num_po_mean_pc
    ) %>% 
    kable(booktabs = T,
          col.name = c("Question category",
                       "M", "SD",
                       "M", "% of category",
                       "M", "% of category")) %>% 
    kable_styling(full_width = T) %>%
    add_header_above(c(" " = 1, "Questions per lecture" = 2, "Student responses per lecture" = 2, "Participation opportunities per lecture" = 2))
```

```{r echo=FALSE}
counts_by_lecture_summary_table_noAU %>% 
    select(
      code_name,
      num_q_mean,
      num_q_sd,
      num_sr_mean,
      num_sr_mean_pc,
      num_po_mean,
      num_po_mean_pc
    ) %>% 
    kable(booktabs = T,
          format = "latex",
          col.name = c("Question category",
                       "M", "SD",
                       "M", "% of category",
                       "M", "% of category")) %>% 
    kable_styling(full_width = T) %>%
    add_header_above(c(" " = 1, "Questions per lecture" = 2, "Student responses per lecture" = 2, "Participation opportunities per lecture" = 2)) %>% 
  cat()
```


```{r}
plot_summary(counts_by_lecture_inc_totals_noAU)
```

#### Session info

```{r}
sessionInfo()
```