---
title: "Lecturer Questions: Wait times"
author: "George Kinnear"
date: "2024-04-04"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
theme_set(theme_minimal())
```

```{r read-data, message=FALSE, warning=FALSE}
all_qs_all_lecs <- read_csv("data/ANON_LQ_Maths_All.csv")
anon_courses <- read_csv("data/ANON_courses.csv")

codes <- c("A", "C", "E", "F", "N", "O", "P", "U", "W", "X")
# Collapsing C, P, X, O into Other
collapsed_codes <- c("A", "O", "E", "F", "N", "O", "O", "U", "W", "O")

code_names <- c("Admin", "Convention", "Evaluation", "Fact", "Next step", "Other",
                "Proof framework", "Understanding check", "Warrant", "Example generation")

code_details <- tibble(code = codes, code_name = code_names) %>% 
  mutate(
    collapsed_code = case_when(
      code %in% c("C", "P", "X", "O") ~ "Other",
      TRUE ~ code_name
    )
  ) %>% 
  add_row(
    code = "Total",
    code_name = "Total",
    collapsed_code = "Total"
  )

# Making list of all lectures and category code combos
lec_cat <- expand_grid(
  all_qs_all_lecs %>% select(course, session) %>% distinct(),
  category = codes
)
```

Just as a check, here is the total number of mathematical questions in the data we have loaded:

```{r}
all_qs_all_lecs %>% 
  filter(!is.na(code_name)) %>% 
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  tally(name = "num_qs") %>%
  basic_kable()
```

# Wait time for each question

Wait times are binned into the same ranges as in Paoletti et al. (2018):

* 0
* 1-2 seconds
* 3-4 seconds
* 5+ seconds

```{r}
qs_and_wait_time <- all_qs_all_lecs %>%
  select(course, session, code_name, wait_time) %>%
  filter(!is.na(code_name)) %>% 
  mutate(
    wait_time_bin = case_when(
      wait_time == 0 ~ "0",
      wait_time <3 ~ "1-2",
      wait_time <5 ~ "3-4",
      wait_time >4 ~ "5+"
    )
  )
```

Here we show the wait time for every (mathematical) question in the sample, grouped by question type and with the time bins highlighted:

```{r}
qs_and_wait_time %>% 
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  ggplot(aes(x = code_name)) +
  geom_jitter(aes(y = wait_time, colour = wait_time_bin), alpha = 0.3, size = 2, stroke = 0) +
  #geom_point(aes(y = num_q_mean, colour = course_level), alpha = 1, size = 3) +
  stat_summary(aes(y = wait_time), fun=mean, geom="point", shape=20, size=6, colour = "red", alpha = 1, position = position_nudge(x = 0))+
  scale_colour_viridis_d("Wait time (s)", end = 0.8, option = "plasma", direction = -1) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom"
  ) +
  labs(
    x = "Question Type",
    y = "Wait time in seconds"
  ) +
  coord_flip()
ggsave("figs/FIG_Wait_times.pdf",width=30,height=15,units="cm",dpi=300)
```

## Summary statistics

This shows the mean (and SD) of the number of questions of each type and wait time **per lecture**.

```{r}

qs_and_wait_time_completed <- qs_and_wait_time %>% 
  group_by(wait_time_bin, course, session, code_name) %>% 
  summarise(
    num_qs = n(),
    .groups = "drop"
  ) %>% 
  # make sure lectures with no questions are included, with counts of 0
  right_join(
    lec_cat %>%
      left_join(code_details, by = c("category" = "code")) %>% 
      select(course, session, code_name) %>% 
      expand_grid(
        qs_and_wait_time %>% select(wait_time_bin) %>% distinct()
      ),
    by = c("wait_time_bin", "course", "session", "code_name")
  ) %>% 
  replace_na(list(num_qs = 0)) %>% 
  filter(!code_name %in% c("Admin", "Understanding check"))

wait_times_summary <- qs_and_wait_time_completed %>% 
  group_by(wait_time_bin, code_name) %>% 
  summarise(
    n = n(),
    sum = sum(num_qs),
    num_qs_mean = mean(num_qs),
    num_qs_sd = sd(num_qs),
    .groups = "drop"
  )

wait_times_summary %>% 
  mutate(
    mean_and_sd = case_when(
      is.na(num_qs_sd) ~ paste0(round(num_qs_mean, digits = 2), " "),
      TRUE ~ paste0(round(num_qs_mean, digits = 2), " (", round(num_qs_sd, digits = 2),")")
    )
  ) %>% 
  select(code_name, wait_time_bin, mean_and_sd) %>% 
  pivot_wider(
    names_from = wait_time_bin,
    values_from = mean_and_sd
  ) %>%  
  kable(booktabs = T, caption = "Mean (sd)") %>% 
  kable_styling(full_width = F)
```

Proportion of all questions with given wait time:

```{r}
# Totals
wait_times_summary_totalsrow <- qs_and_wait_time_completed %>% 
  # add up all the questions of any type in a given session
  group_by(wait_time_bin, course, session) %>% 
  summarise(
    num_qs = sum(num_qs),
    .groups = "drop"
  ) %>%
  group_by(wait_time_bin) %>% 
  summarise(
    n = n(),
    num_qs_mean = mean(num_qs),
    num_qs_sd = sd(num_qs),
    .groups = "drop"
  ) %>% 
  ungroup() %>% 
  mutate(code_name = "Total")

wait_times_summary_totalsrow %>%
  adorn_totals() %>% 
  mutate(
    num_qs_prop = num_qs_mean / sum(num_qs_mean)*2,
    cell_entry = paste0(round(num_qs_mean, digits = 2), " (", round(num_qs_prop*100, digits = 0),"%)")
  ) %>% 
  select(wait_time_bin, cell_entry) %>%
  basic_kable(booktabs = T)
```

This **per lecture** summary is analagous to Table 5 of Paoletti et al. (2018):

```{r message=FALSE, warning=FALSE}
counts_by_lecture_inc_totals_noAU <- read_csv("data/output-03-counts_by_lecture_inc_totals_noAU.csv")
```


```{r}
wait_times_summary_wide_table <- wait_times_summary %>% 
  bind_rows(
    wait_times_summary_totalsrow
  ) %>% 
  select(
    wait_time_bin,
    code_name,
    mean = num_qs_mean,
    sd = num_qs_sd
  ) %>% 
  # add percentages of category
  group_by(code_name) %>% 
  mutate(
    type_pc = round(mean / sum(mean) * 100, 0)
  ) %>% 
  ungroup() %>% 
  pivot_wider(
    names_from = wait_time_bin,
    values_from = c(mean, sd, type_pc),
    names_glue = "{wait_time_bin}_{.value}"
  ) %>% 
  # add the totals for each category, from earlier table
  left_join(
    counts_by_lecture_inc_totals_noAU %>% group_by(code_name) %>% summarise(num_q_mean = mean(num_q)),
    by = "code_name"
  ) %>% 
  arrange(if_else(code_name=="Total",1,0),-num_q_mean) %>% 
  # sort the columns in alphabetical order
  select(code_name, num_q_mean, sort(tidyselect::peek_vars()))

options(knitr.kable.NA = '-')
wait_times_summary_wide_table %>% 
    kable(booktabs = T,
          digits = 2,
          col.name = c("Question category",
                       "M",
                       rep(c("M", "SD", "% of cat"), 4))
    ) %>% 
    kable_styling(full_width = T) %>%
    add_header_above(c(" " = 2, "0s" = 3, "1-2s" = 3, "3-4s" = 3, "5+s" = 3))
```
<details><summary>LaTeX table</summary>
```{r}
wait_times_summary_wide_table %>% 
    kable(booktabs = T,
          format = "latex",
          digits = 2,
          col.name = c("Question category",
                       "M",
                       rep(c("M", "SD", "% of cat"), 4))
    ) %>% 
    kable_styling(full_width = T) %>%
    add_header_above(c(" " = 2, "0s" = 3, "1-2s" = 3, "3-4s" = 3, "5+s" = 3)) %>% 
  cat()
```
</details>

## As proportions of the overall number of questions

Those per lecture numbers are quite small and hard to interpret.

Instead, we work out the proportions more simply, as the proportion of overall number of questions. It gives the same percentages across the rows as the method above, e.g. 23% of Fact questions have a wait time of 0s.

```{r}
qs_and_wait_time_simple_summary <- qs_and_wait_time %>% 
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  tabyl(code_name, wait_time_bin) %>% 
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns(position = "front") %>%
  relocate(Total, .after = code_name) %>% 
  # remove the percentage from the Total column, and sort by it
  mutate(Total = parse_number(str_replace(Total, " \\(.*", ""))) %>% 
  arrange(code_name=="Total", -Total) %>% 
  # Tidy up the column names
  rename("Question category" = code_name) %>%
  rename_with(~ paste(.x, "s (%)"), .cols = matches("\\d"))

qs_and_wait_time_simple_summary %>% 
    basic_kable(booktabs = T, caption = "[Table 7 in the manuscript]. Number of questions of each type broken down by wait time.") %>% 
    kable_styling(full_width = T)
```

<details><summary>LaTeX table</summary>
```{r}
qs_and_wait_time_simple_summary %>% 
  kable(booktabs = T, format = "latex") %>% 
  cat()
```
</details>


## Comparison with Paoletti et al. (2018)

Here we restrict to the courses that were deemed most comparable to those in the Paoletti et al. (2018) sample.

The proportions here are not markedly different from those in the full sample.

```{r}
adv_qs_and_wait_time_simple_summary <- qs_and_wait_time %>% 
  # restrict to only those courses that were deemed "advanced mathematics"
  semi_join(
    anon_courses %>% filter(advanced == "yes"),
    by = c("course" = "anon_code")
  ) %>% 
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  tabyl(code_name, wait_time_bin) %>% 
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns(position = "front") %>%
  relocate(Total, .after = code_name) %>% 
  # remove the percentage from the Total column, and sort by it
  mutate(Total = parse_number(str_replace(Total, " \\(.*", ""))) %>% 
  arrange(code_name=="Total", -Total) %>% 
  # Tidy up the column names
  rename("Question category" = code_name) %>%
  rename_with(~ paste(.x, "s (%)"), .cols = matches("\\d"))

adv_qs_and_wait_time_simple_summary %>% 
    basic_kable(booktabs = T, caption = "For the proof-based courses only, number of questions of each type broken down by wait time.") %>% 
    kable_styling(full_width = T)
```

<details><summary>LaTeX table</summary>
```{r}
adv_qs_and_wait_time_simple_summary %>% 
  kable(booktabs = T, format = "latex") %>% 
  cat()
```
</details>


Going even further, if we get rid of 1C which dominates the sample, the wait times skew higher:

```{r}
non1c_adv_qs_and_wait_time <- qs_and_wait_time %>% 
  semi_join(
    # restrict to only those courses that were deemed "advanced mathematics"
    anon_courses %>%
      filter(advanced == "yes") %>%
      # and further restrict to avoid the course 1C which has many lectures in the sample
      filter(anon_course != "1C"),
    by = c("course" = "anon_code")
  ) %>% 
  filter(!code_name %in% c("Admin", "Understanding check"))

non1c_adv_qs_and_wait_time_simple_summary <- non1c_adv_qs_and_wait_time %>% 
  tabyl(code_name, wait_time_bin) %>% 
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns(position = "front") %>%
  relocate(Total, .after = code_name) %>% 
  # remove the percentage from the Total column, and sort by it
  mutate(Total = parse_number(str_replace(Total, " \\(.*", ""))) %>% 
  arrange(code_name=="Total", -Total) %>% 
  # Tidy up the column names
  rename("Question category" = code_name) %>%
  rename_with(~ paste(.x, "s (%)"), .cols = matches("\\d"))

non1c_adv_qs_and_wait_time_simple_summary %>% 
    basic_kable(booktabs = T, caption = "For the proof-based courses only, and excluding course 1C, number of questions of each type broken down by wait time.") %>% 
    kable_styling(full_width = T)
```

For reference, this subset of the sample has `r non1c_adv_qs_and_wait_time %>% distinct(course, session) %>% nrow()` lectures.

## Wait time and responses

```{r}
wait_time_vs_response_tab <- all_qs_all_lecs %>%
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  filter(!is.na(code_name)) %>% 
  mutate(
    wait_time_bin = case_when(
      wait_time == 0 ~ "0",
      wait_time <3 ~ "1-2",
      wait_time <5 ~ "3-4",
      wait_time >4 ~ "5+"
    )
  ) %>% 
  count(wait_time_bin, student_response)

wait_time_vs_response_tab %>% 
  group_by(wait_time_bin) %>% 
  mutate(pct_label = str_glue("{n}\n({round(n / sum(n) * 100, 0)}%)")) %>% 
  ggplot(aes(x = wait_time_bin, y = n, fill = student_response)) +
    geom_bar(position = "fill", stat = "identity") +
    geom_text(aes(label = pct_label), position = position_fill(vjust = 0.5), colour = "white") +
    scale_fill_viridis_d("Student response", begin = 0, end = 0.7, option = "plasma") +
    labs(x = "Wait time", y = "Proportion of questions")

wait_time_vs_response_tab %>% 
  pivot_wider(names_from = "student_response", values_from = "n") %>% 
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns(position = "front") %>% 
  relocate(Total, .after = wait_time_bin) %>% 
  mutate(Total = parse_number(str_replace(Total, " \\(.*", ""))) %>% 
  basic_kable(
    booktabs = T,
    caption = "Response rates for questions by wait time.",
    col.names = c("Wait time", "Total", "No", "Yes")
  ) %>% 
  kable_styling(full_width = T) %>%
  add_header_above(c(" " = 2, "Student response" = 2))
```
Considering only questions not part of a chain:

```{r}
wait_time_vs_response_tab_nochain <- all_qs_all_lecs %>%
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  filter(!is.na(code_name)) %>% 
  mutate(
    wait_time_bin = case_when(
      wait_time == 0 ~ "0",
      wait_time <3 ~ "1-2",
      wait_time <5 ~ "3-4",
      wait_time >4 ~ "5+"
    )
  ) %>% 
  filter(cluster_q == "N") %>% 
  count(wait_time_bin, student_response)

wait_time_vs_response_tab_nochain %>% 
  group_by(wait_time_bin) %>% 
  mutate(pct_label = str_glue("{n}\n({round(n / sum(n) * 100, 0)}%)")) %>% 
  ggplot(aes(x = wait_time_bin, y = n, fill = student_response)) +
    geom_bar(position = "fill", stat = "identity") +
    geom_text(aes(label = pct_label), position = position_fill(vjust = 0.5), colour = "white") +
    scale_fill_viridis_d("Student response", begin = 0, end = 0.7, option = "plasma") +
    labs(x = "Wait time", y = "Proportion of questions", title = "Response to questions not part of a chain")

wait_time_vs_response_tab_nochain %>% 
  pivot_wider(names_from = "student_response", values_from = "n") %>% 
  adorn_totals(c("row", "col")) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns(position = "front") %>% 
  relocate(Total, .after = wait_time_bin) %>% 
  mutate(Total = parse_number(str_replace(Total, " \\(.*", ""))) %>% 
  basic_kable(
    booktabs = T,
    caption = "Response rates for questions by wait time.",
    col.names = c("Wait time", "Total", "No", "Yes")
  ) %>% 
  kable_styling(full_width = T) %>%
  add_header_above(c(" " = 2, "Student response" = 2))
```

What about the position of the question in a chain?

```{r}
wait_time_by_chain_position <- all_qs_all_lecs %>%
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  filter(!is.na(code_name)) %>% 
  mutate(
    wait_time_bin = case_when(
      wait_time == 0 ~ "0",
      wait_time <3 ~ "1-2",
      wait_time <5 ~ "3-4",
      wait_time >4 ~ "5+"
    )
  ) %>% 
  group_by(course, session, cluster_q) %>% 
  mutate(
    chain_status = case_when(
      cluster_q == "N" ~ "Single question",
      row_number() == max(row_number()) ~ "End of chain",
      .default = "Mid chain"
    )
  ) %>% 
  group_by(chain_status, wait_time_bin, student_response) %>% 
  tally()

wait_time_by_chain_position %>% 
  group_by(chain_status, wait_time_bin) %>% 
  mutate(pct_label = str_glue("{n}\n({round(n / sum(n) * 100, 0)}%)")) %>% 
  ggplot(aes(x = wait_time_bin, y = n, fill = student_response)) +
    geom_bar(position = "fill", stat = "identity") +
    geom_text(aes(label = pct_label), position = position_fill(vjust = 0.5), colour = "white", size = 3) +
    scale_fill_viridis_d("Student response", begin = 0, end = 0.7, option = "plasma") +
    labs(x = "Wait time", y = "Proportion of questions", title = "Response to questions not part of a chain") +
  facet_wrap(~ chain_status)

```

<details><summary>Table of values</summary>

```{r}
wait_time_by_chain_position %>% 
  group_by(chain_status, wait_time_bin) %>% 
  mutate(pct_label = str_glue("{round(n / sum(n) * 100, 0)}%")) %>% 
  basic_kable()
```

</details>



### Response to questions with long wait time

```{r}
long_wait_time_single_qs <- all_qs_all_lecs %>%
  filter(!code_name %in% c("Admin", "Understanding check")) %>% 
  filter(!is.na(code_name)) %>% 
  filter(wait_time > 4) %>% 
  filter(cluster_q == "N")
```


These are different perspectives on the `r nrow(long_wait_time_single_qs)` questions that were not part of a chain, and that had wait time of 5+ seconds.

```{r}
long_wait_time_single_qs %>% 
  ggplot(aes(x = student_response, y = wait_time)) +
  geom_boxplot()

long_wait_time_single_qs %>% 
  group_by(category = collapsed_code, student_response) %>% 
  tally() %>% 
  group_by(student_response) %>% 
  mutate(bar_width = sum(n)) %>% 
  ungroup() %>%
  mutate(
    category = fct_relevel(as_factor(category), "Fact", "Next step", "Warrant", "Evaluation", "Other")
  ) %>%
  mutate(bar_width = bar_width / sum(n)) %>% 
  ggplot(aes(x = student_response)) +
  geom_bar(aes(y = n, fill = category, width = bar_width), position = "fill", stat = "identity") +
  scale_fill_viridis_d("Category", begin = 0.1, end = 0.8, option = "plasma", direction = -1) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
  ) +
  labs(
    x = "Question got a response",
    y = "Proportion of questions"
  )

long_wait_time_single_qs %>% 
  group_by(category = collapsed_code, student_response) %>% 
  tally() %>% 
  ggplot(aes(x = category)) +
  geom_bar(aes(y = n, fill = student_response), position = "fill", stat = "identity") +
  scale_fill_viridis_d("Question got a response", begin = 0.1, end = 0.8, option = "plasma", direction = -1) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
  ) +
  labs(
    x = "Question category",
    y = "Proportion of questions"
  )
```



<details><summary>Table with question details</summary>

```{r}
long_wait_time_single_qs %>% 
  select(course, session, student_response, code_name, question, wait_time) %>% 
  arrange(wait_time) %>% 
  basic_kable()
```

</details>



<details><summary>Session info</summary>
```{r}
sessionInfo()
```
</details>