---
title: "Lecturer Questions: Question chains"
author: "George Kinnear"
date: "17/11/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
theme_set(theme_minimal())
```

```{r read-data, message=FALSE, warning=FALSE}
all_qs_all_lecs <- read_csv("data/ANON_LQ_Maths_All.csv")
anon_courses <- read_csv("data/ANON_courses.csv")

codes <- c("A", "C", "E", "F", "N", "O", "P", "U", "W", "X")
# Collapsing C, P, X, O into Other
collapsed_codes <- c("A", "O", "E", "F", "N", "O", "O", "U", "W", "O")

code_names <- c("Admin", "Convention", "Evaluation", "Fact", "Next step", "Other",
                "Proof framework", "Understanding check", "Warrant", "Example generation")

code_details <- tibble(code = codes, code_name = code_names) %>% 
  mutate(
    collapsed_code = case_when(
      code %in% c("C", "P", "X", "O") ~ "Other",
      TRUE ~ code_name
    )
  ) %>% 
  add_row(
    code = "Total",
    code_name = "Total",
    collapsed_code = "Total"
  )

# Making list of all lectures and category code combos
lec_cat <- expand_grid(
  all_qs_all_lecs %>% select(course, session) %>% distinct(),
  category = codes
)
```

Just as a check, here is the total number of questions in the data we have loaded:

```{r}
all_qs <- all_qs_all_lecs %>% 
  filter(!is.na(code_name))

all_qs %>% 
  tally(name = "num_qs") %>%
  basic_kable()
```

# Question chains


## Prevalence

A majority of questions are part of a chain:

```{r}
all_qs %>% 
  mutate(part_of_chain = str_detect(cluster_q, "Y")) %>% 
  tabyl(part_of_chain) %>%
  adorn_pct_formatting() %>% 
  basic_kable(booktabs = T)
```


## Untangling U and A

How often do U and U questions crop up in chains? Ideally we would like to filter those out before the analyses, but that may not make sense if they are highly prevalent in chains

```{r}
numbered_chains <- all_qs %>% 
  #filter(!category %in% c("A", "U")) %>% 
  filter(str_detect(cluster_q, "Y")) %>% 
  group_by(course, session, cluster_q) %>% 
  mutate(
    chain_id = cur_group_id(),
    num_in_chain = n(),
    pos_within_chain = row_number(),
    excluded_qs = sum(category=="A") + sum(category=="U")
  ) %>% 
  bind_rows(
    all_qs %>% 
      filter(cluster_q == "N") %>% 
      mutate(num_in_chain = 1, pos_within_chain = 1, excluded_qs = 1)
  ) %>% 
  mutate(num_in_chain = as.numeric(num_in_chain))

chains_excluding_AU_pivot <- numbered_chains %>% 
  ungroup() %>% 
  # add a chain ID for each single question
  mutate(
    chain_id = ifelse(is.na(chain_id), 5000+row_number(), chain_id),
  ) %>% 
  select(chain_id, num_in_chain, excluded_qs) %>% 
  distinct() %>% 
  tabyl(num_in_chain, excluded_qs)

chains_excluding_AU_pivot %>% 
  basic_kable(booktabs = T) %>% 
  add_header_above(c(" " = 1, "Number of A/U questions excluded" = 5))
```

This shows that in almost all cases, the number of excluded questions matches the size of the chain -- so chains tend to be either all A/U or all non-A/U.

Investigating this further -- here we pick out all the cases where a chain is partly made up of A/U:

```{r}
numbered_chains %>% 
  ungroup() %>% 
  filter(excluded_qs > 0, excluded_qs < num_in_chain) %>% 
  select(course, session, chain_id, cluster_q, category, question) %>% 
  basic_kable(booktabs = T,
        caption = "Chains where one of the questions is A/U") %>%
  collapse_rows(columns = 1:4, valign = "top")
```

These are really quite rare:

```{r}
affected_chains <- numbered_chains %>% 
  ungroup() %>% 
  mutate(affected = excluded_qs > 0 & excluded_qs < num_in_chain) %>% 
  select(chain_id, affected) %>% 
  distinct() %>% 
  tabyl(affected)

affected_chains %>% 
  basic_kable(booktabs = T)
```

Thus, in the following analyses, we first **omit all A/U questions**.


# Question chain frequencies

After discarding all A/U questions, chains are even more frequent:

```{r}
all_qs %>% 
  filter(!category %in% c("A", "U")) %>% 
  mutate(part_of_chain = str_detect(cluster_q, "Y")) %>% 
  tabyl(part_of_chain) %>% 
  adorn_pct_formatting() %>% 
  basic_kable(booktabs = T)
```


How frequent are chains of different lengths?

```{r}
numbered_chains <- all_qs %>% 
  filter(!category %in% c("A", "U")) %>% 
  filter(str_detect(cluster_q, "Y")) %>% 
  group_by(course, session, cluster_q) %>% 
  mutate(
    chain_id = cur_group_id(),
    num_in_chain = n(),
    pos_within_chain = row_number()
  ) %>% 
  bind_rows(
    all_qs %>% 
      filter(!category %in% c("A", "U")) %>% 
      filter(cluster_q == "N") %>% 
      mutate(num_in_chain = 1, pos_within_chain = 1)
  ) %>% 
  mutate(num_in_chain = as.numeric(num_in_chain)) %>% 
  # remove the very small number of chains (5) which now only have 1 questions after A/U were removed
  mutate(chain_id = ifelse(num_in_chain==1, NA, chain_id)) %>% 
  ungroup()

numbered_chains %>% 
  filter(!is.na(chain_id)) %>% 
  # make sure all lectures appear, including those with counts of 0
  right_join(
    lec_cat %>% select(course, session),
    by = c("course", "session")
  ) %>% 
  replace_na(list(num_in_chain = 0)) %>% 
  group_by(course, session) %>% 
  summarise(
    num_chains = n_distinct(chain_id, na.rm = TRUE),
    chain_size_mean = mean(num_in_chain, na.rm = TRUE),
    chain_size_max = max(num_in_chain, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  head() %>% 
  basic_kable(booktabs = T,
              caption = "Excerpt of table showing number of chains and the mean/max chain size in each session")
```

Looking at the chains, they tend to be quite small but there are a few larger ones:

```{r}
numbered_chains %>% 
  #filter(!is.na(chain_id)) %>% 
  #ungroup() %>%
  filter(num_in_chain > 1) %>% 
  select(chain_id, num_in_chain) %>% 
  distinct() %>% 
  summarise(
    num_chains = n_distinct(chain_id),
    chain_size_mean = mean(num_in_chain),
    chain_size_median = median(num_in_chain),
    chain_size_sd = sd(num_in_chain),
    chain_size_max = max(num_in_chain)
  ) %>% 
  basic_kable()
```

```{r}
numbered_chains %>% 
  filter(!is.na(chain_id)) %>% 
  select(num_in_chain, chain_id) %>% 
  distinct() %>% 
  tabyl(num_in_chain) %>% 
  adorn_pct_formatting() %>% 
  basic_kable()

numbered_chains %>% 
  filter(!is.na(chain_id)) %>% 
  select(num_in_chain, chain_id) %>% 
  distinct() %>% 
  ggplot(aes(x = num_in_chain)) +
    geom_histogram(binwidth = 1)
```



## Chains and responses

```{r}
chain_response_rates <- numbered_chains %>% 
  filter(!category %in% c("A", "U")) %>% 
  ungroup() %>% 
  #filter(!is.na(chain_id)) %>% 
  mutate(
    chain_id = ifelse(is.na(chain_id), 5000+row_number(), chain_id)
  ) %>% 
  group_by(num_in_chain, chain_id) %>% 
  summarise(got_a_response = last(student_response), .groups = "drop") %>% 
  tabyl(num_in_chain, got_a_response) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns()
```


```{r}
chain_responses_table <- numbered_chains %>% 
  #filter(!is.na(chain_id)) %>% 
  mutate(
    chain_id = ifelse(is.na(chain_id), 5000+row_number(), chain_id),
    num_in_chain = ifelse(num_in_chain==1, "Single question", "Question chain")
  ) %>% 
  group_by(num_in_chain, chain_id) %>% 
  summarise(got_a_response = last(student_response), .groups = "drop") %>% 
  tabyl(num_in_chain, got_a_response) %>% 
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 0) %>%
  adorn_ns() %>% 
  arrange(desc(num_in_chain)) %>% 
  # add rows showing the detail on genuine chains
  bind_rows(
    chain_response_rates %>%
      filter(num_in_chain > 1) %>%
      #mutate(num_in_chain = as.character(num_in_chain))
      mutate(num_in_chain = paste(num_in_chain, "questions"))
  )

chain_responses_table %>% 
  kable(booktabs = T,
         col.names = c("", "No", "Yes")) %>% 
    kable_styling(full_width = F) %>%
    #pack_rows("Question chain size", 3, 9) %>% 
    add_indent(c(3:9)) %>% 
    add_header_above(c(" " = 1, "Student response" = 2))
```

<details><summary>LaTeX table</summary>
```{r}
chain_responses_table %>% 
  kable(booktabs = T,
        format = "latex",
         col.names = c("", "No", "Yes")) %>% 
    kable_styling(full_width = F) %>%
    #pack_rows("Question chain size", 3, 9) %>% 
    add_indent(c(3:9)) %>% 
    add_header_above(c(" " = 1, "Student response" = 2)) %>% 
  cat()
```
</details>


# Analysis of the chains


```{r}
chain_summaries <- numbered_chains %>%
  filter(!category %in% c("A", "U")) %>% 
  ungroup() %>% 
  mutate(
    chain_id = ifelse(is.na(chain_id), 5000+row_number(), chain_id)
  ) %>% 
  group_by(chain_id) %>% 
  summarise(
    category_chain = paste0(category, collapse = ''),
    chain_len = n(),
    codes_used = n_distinct(category)
  ) %>% 
  arrange(-chain_len)
```

Most chains consist of questions of a single type:

```{r}
options(knitr.kable.NA = '')

chain_summaries_nums <-
chain_summaries %>% 
  group_by(chain_len, codes_used) %>% 
  tally() %>% 
  pivot_wider(
    names_from = codes_used,
    values_from = n
  )

chain_summaries_nums %>% 
  filter(chain_len > 1) %>% 
  # make the chain_len a string so it's not added to the totals!
  mutate(chain_len = paste("", chain_len)) %>% 
  adorn_totals() %>% 
  basic_kable() %>%
  add_header_above(c(" " = 1, "Number of codes used" = 3))
```


Look at the more diverse chains:

```{r}
diverse_chains <- 
chain_summaries %>% 
  filter(codes_used > 1) %>% 
  filter(codes_used > 2 | chain_len+codes_used > 5) %>% 
  arrange(-codes_used, chain_len)

numbered_chains %>% 
  ungroup() %>% 
  right_join(diverse_chains, by = "chain_id") %>% 
  select(category_chain, course, session, category, question)
```

These are all the chains of size 2 where both questions are of the same type:

```{r}
numbered_chains %>% 
  ungroup() %>% 
  right_join(
    chain_summaries %>% 
      filter(codes_used == 1, chain_len == 2),
    by = "chain_id"
  ) %>% 
  select(chain_id, category_chain, course, session, category, question)
```
This shows the number of chains of each type signature:

```{r}
chain_summaries %>% 
  filter(chain_len > 1) %>%
  group_by(category_chain) %>%
  tally() %>% 
  arrange(-n) %>% 
  basic_kable()
```



<details><summary>Session info</summary>
```{r}
sessionInfo()
```
</details>